Automatic Matching of Bullet Lands

### 2016-01-17 imaging-paper.Rnw
typo

Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
\hh{XXX Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}. There, CMS is defined only in terms of matching peaks. Additionally, peaks in  \citet{thompson:2013} are only used if they can be identified and matched `within a tolerable range' between lands. The definition given here is computationally less complex, but should yield similar values, because of the requirement to only consider signatures from a stable region in the land (see Section~\ref{sec:heights} for further details on stability of regions). In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92). XXX}
For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match. \hh{XXX Biasotti's four or more consecutive lines is closer to eight or more with our definition. Should we just multiply our CMS by 0.5, to account for the fact that between every two peaks has to be a valley? XXX} We will assess this claim in the next section.

@@

set.seed(20140105)
aucs <- plyr::ldply(features, function(x) {
  subbstats <- subset(bstats, flagged == FALSE)
  pos.scores <- sample(subbstats[which(subbstats$match),x], 10000, replace=TRUE)
  neg.scores <- sample(subbstats[which(!subbstats$match),x], 10000, replace=TRUE)
  data.frame(x, upper=mean(pos.scores > neg.scores), 
             lower = mean(pos.scores < neg.scores))
})
aucs <- melt(aucs, measure.var=c("lower", "upper"))
names(aucs) <- c("variable", "type", "auc")
rocs <- merge(rocs, aucs, by=c("variable", "type"))
rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")

@@

Besides resulting in a probabilistic quantification of matches, random forests also provide an assessment of the importance of each of the features derived from the bullets' 3D topological surface measurements. Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to supplement section~\ref{supp:randomforest}). 
\begin{figure}
<<echo=FALSE, fig.width=7, fig.height=2, out.width='.7\\textwidth'>>=
imp <- data.frame(importance(rtrees))
imp$Variable <- row.names(imp)
imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
imp1 <- subset(imp, !(Variable %in% c("x1", "sd.D")))
imp2 <- subset(imp, Variable %in% c("x1", "sd.D"))
qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
  geom_text(aes(y=1.2, label=Variable), angle=45, hjust=0, data=imp1) + 
  geom_text(aes(y=0.8, label=Variable), angle=-45, hjust=0, data=imp2) + 
  ylim(c(0, 2.75)) + theme_bw() + ylab("") +
  theme(axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        panel.grid.major.y=element_blank(), panel.grid.minor.y=element_blank()) +
  xlab("Importance (mean decrease in Gini index)")
@
\caption{\label{fig:importance}Importance of features in the random forest. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.}
\end{figure}

@@

\hh{Besides including results from known matches againts known non-matches, we can increase the number of comparisons in the Hamby study to include all possible land-to-land comparisons. This effectively doubles the number of data points available. Comparisons not previously included in fitting the random forest can also be used as an additional source for assessing error rates.  Results for this and a more detailed discussion can be found in supplement section~\ref{supp:extended}.}

>>>>

Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
\hh{ Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}. There, CMS is defined only in terms of matching peaks without regarding valleys. Additionally, peaks in  \citet{thompson:2013} are  used only if they can be identified and matched `within a tolerable range' between lands. The definition given here is computationally less complex, but should yield highly correlated values, because of the requirement to only consider signatures from a stable region in the land (see Section~\ref{sec:heights} for further details on stability of regions). In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92). }
For lead bullets, such as used in the Hamby study, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match. \hh{Biasotti's four or more consecutive lines translates to eight or more with our definition. } %We will assess the claim that eight or more consecutive lines are sufficient for a match in the next section.

@@

set.seed(20140105)
aucs <- plyr::ldply(features, function(x) {
  subbstats <- subset(bstats, flagged == FALSE)
  pos.scores <- sample(subbstats[which(subbstats$match),x], 50000, replace=TRUE)
  neg.scores <- sample(subbstats[which(!subbstats$match),x], 50000, replace=TRUE)
  data.frame(x, upper=mean(pos.scores > neg.scores), 
             lower = mean(pos.scores < neg.scores))
})
aucs <- melt(aucs, measure.var=c("lower", "upper"))
names(aucs) <- c("variable", "type", "auc")
rocs <- merge(rocs, aucs, by=c("variable", "type"))
rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")

@@

Besides resulting in a probabilistic quantification of matches, random forests also provide an assessment of the importance of each of the features derived from the bullets' 3D topological surface measurements. Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to supplement section~\ref{supp:randomforest}). 
\begin{figure}
<<echo=FALSE, fig.width=7, fig.height=2, out.width='.7\\textwidth'>>=
imp <- data.frame(importance(rtrees))
imp$Variable <- row.names(imp)
imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
imp1 <- subset(imp, !(Variable %in% c("x1", "sd.D", "S")))
imp2 <- subset(imp, Variable %in% c("x1", "sd.D", "S"))
qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
  geom_text(aes(y=1.2, label=Variable), angle=45, hjust=0, data=imp1) + 
  geom_text(aes(y=0.8, label=Variable), angle=-45, hjust=0, data=imp2) + 
  ylim(c(0, 2.75)) + theme_bw() + ylab("") +
  theme(axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        panel.grid.major.y=element_blank(), panel.grid.minor.y=element_blank()) +
  xlab("Importance (mean decrease in Gini index)")
@
\caption{\label{fig:importance}Importance of features in the random forest. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.}
\end{figure}

@@

\hh{Besides including results from known matches against known non-matches, we can increase the number of comparisons in the Hamby study to include all possible land-to-land comparisons. This effectively doubles the number of data points available. Comparisons not previously included in fitting the random forest can also be used as an additional source for assessing error rates.  Results for this and a more detailed discussion can be found in supplement section~\ref{supp:extended}.}
